Optimal strategies of influence in structured populations

### 2015-03-19 paper/manuscript.tex
start of PRL write-up

\documentclass[3p,number,sort&compress,times]{elsarticle}
\usepackage{amsmath,amsfonts}  %for mathstyles
\usepackage{hyperref}
\usepackage{color}
%\usepackage{colortbl}
\usepackage{wasysym}
%\usepackage{ifsym}
\usepackage{array}
\usepackage{longtable}
\usepackage{graphicx}
\journal{..}
%opening

\begin{document}
\begin{frontmatter}
%\preprint{}
\title{Optimal strategies of influence in structured populations}
\author[ex]{Colin J. Torney}
\author[eeb]{Simon A. Levin}
%
%\cortext[cor1]{Corresponding author}
\address[ex]{Centre for Mathematics \& the Environment, University of Exeter, Penryn Campus, Cornwall, UK}
\address[eeb]{Department of Ecology \& Evolutionary Biology, Princeton University, Princeton, NJ, USA}

\begin{abstract}
Conventions, opinions, or social norms often persist within interacting populations long after their usefulness has ended or their inaccuracy revealed. In many scenarios this is due to conformity by individuals, or the devaluation of individual preferences and opinions when they contradict observed social information. Both the behavior at the individual-level and the structure of the social network play a role in transitions away from out-dated, collectively-held opinions. Personal preferences, biases and the predilection to conform are significant factors, however system level characteristics, such as the statistical properties of the network, or the rate of social mixing, are also strong determinants of macroscale behavior. When seeking to influence a collective decision, different strategies may be more or less effective depending on these factors. Here we show that the effectiveness of a given strategy of influence is dependent on the nature of the underlying interaction network. 
\end{abstract}

\end{frontmatter}

@@

We aim to assess optimal strategies assuming some cost associated with reducing the noise. Cost is per unit time and assumed to scale exponentially so that forcing $\sigma \rightarrow 0$ incurs exponentially increasing costs,
\begin{equation}
\mbox{Cost} = \tau (\sigma) e^{\beta N/\sigma}
\label{cost}
\end{equation}
Minimizing Eqn.~\ref{cost} means the optimal level of $\sigma$ is one for which
\begin{equation}
N^{-1} \frac{d \ln \tau}{d \sigma} = \beta \sigma^{-2}
\end{equation}
\begin{figure*}
\centering
\includegraphics[scale=0.4]{figures/fig-chi.eps} 
\caption{Probability to transition as a function of local neighbourhood for different values of the sociality weighting $\chi$}
\label{fig1}
\end{figure*}
We need to find $\tau$.
\section{Mean first passage time away from incorrect state}
Consider the process as a one-step Markov chain on the integer line from $0$ to $N$. A transition from state $j$ to $j+1$ is written $t^+(j)$ and an opposite transition is denoted $t^-(j)$. An exact expression for the mean first passage time from zero individuals correct to $N_{ESC}$ is \cite{gardiner}

@@

\begin{equation}
N \int_{0}^{x_{ESC}} \exp\left[N\phi(x)\right] f(x) dx \approx \sqrt{\frac{2\pi N}{\vert \phi''(x_2) \vert}}  \exp\left[N\phi(x_2)\right] f(x_2) 
\label{laplace_out}
\end{equation}
As $x_1<x_2$ this may be repeated for the inner integral to leave
\begin{equation}
\tau \approx  \frac{2\pi \exp\left[N \phi(x_2)-N \phi(x_1) \right] }{t^+(x_1)\sqrt{\vert \phi''(x_2) \vert \phi''(x_1)}}.
\label{laplace_approx}
\end{equation}
Taking logarithms, derivatives, and again assuming large $N$, leaves
\begin{equation}
N^{-1}  \frac{d \ln \tau}{d \sigma} \approx -\int_{x_1}^{x_2} \frac{\partial}{\partial \sigma}\ln\left(\frac{t^+(y)}{t^-(y)}\right) dy
\label{laplace_approx}
\end{equation}

\section{Transition rates}
The dynamics of an update step in stochastic decision process are as follows; first an individual is selected at random from the population, this individual has a personal estimate of the environmental cue $g_i$ and observes the opinions of all its neighbors. The decision rule of Eqn.~\ref{opt_rule} is then employed and the individual will adopt, or remain in, the state most likely to be correct.
For any value of social information $x_L$ there is a value of personal information above which the individual will switch to $S=1$. We denote this value as $g_c$ and from Eqn.~\ref{opt_rule} 
\begin{equation}
g_c(x_L) = \sigma\ln\chi  \left(\frac{1}{2}-x_L\right) 
\end{equation}
If $g_i>g_c$ then the individual will switch to $S=1$ otherwise $S=-1$. 
Dealing first with the transition rate toward the correct state, we may write this as
\begin{equation}
%t^+(x) = P(S_i=-1)\sum_{j=0}^{N_s}P\left(g_i>g_c(j/N_s)\right)P\left(x_L=j/N_s|x\right)
t^+(x) = P(S_i=-1)\sum_{x_{Li}}P\left(g_i>g_c(x_{Li})\right)P\left(x_{Li}|x\right)
\end{equation}
where $i$ denotes the selected individual and the summation is taken over all possible values of $x_{Li}$.
he product of the probability that the focal individual is in the state $S=-1$ by the probability of a given value of $x_L$ by the

>>>>

\documentclass[aps,prl,twocolumn,showpacs,superscriptaddress,groupedaddress]{revtex4} 
\usepackage{amsmath,amsfonts}  %for mathstyles
\usepackage{hyperref}
\usepackage{color}
%\usepackage{colortbl}
\usepackage{wasysym}
%\usepackage{ifsym}
\usepackage{array}
\usepackage{longtable}
\usepackage{graphicx}
%\journal{..}
%opening

\begin{document}
%\begin{frontmatter}
%\preprint{}
\title{Optimal strategies of influence in structured populations}
\author{Colin J. Torney}
\affiliation{Centre for Mathematics \& the Environment, University of Exeter, Penryn Campus, Cornwall, UK}
%\affiliation{Department of Ecology \& Evolutionary Biology, Princeton University, \\Princeton, New Jersey 08544}
\author{Simon A. Levin}
\affiliation{Department of Ecology \& Evolutionary Biology, Princeton University, \\Princeton, New Jersey 08544}

\begin{abstract}
Incorrect opinions often persist for long periods within interacting populations. In many scenarios this is due to conformity and the devaluation of individual opinions when they contradict observed social information. Both individual characteristics and system-level properties control the expected time to transition away from an out-dated collective opinion. These factors also determine the effectiveness of strategies to engineer transitions to the optimal state. By analysing a model of opinion dynamics, we show here how the optimal strategy of influence relates to the underlying interaction network. 
\end{abstract}
\maketitle
%\end{frontmatter}

@@

We aim to assess optimal strategies assuming some cost associated with reducing the noise. Cost is per unit time and assumed to scale exponentially so that forcing $\sigma \rightarrow 0$ incurs exponentially increasing costs,
\begin{equation}
\mbox{Cost} = \tau (\sigma) e^{\beta N/\sigma}
\label{cost}
\end{equation}
Minimizing Eqn.~\ref{cost} means the optimal level of $\sigma$ is one for which
\begin{equation}
N^{-1} \frac{d \ln \tau}{d \sigma} = \beta \sigma^{-2}
\end{equation}
%\begin{figure*}
%\centering
%\includegraphics[scale=0.4]{figures/fig-chi.eps} 
%\caption{Probability to transition as a function of local neighbourhood for different values of the sociality weighting $\chi$}
%\label{fig1}
%\end{figure*}
We need to find $\tau$.
%\section{Mean first passage time away from incorrect state}
Consider the process as a one-step Markov chain on the integer line from $0$ to $N$. A transition from state $j$ to $j+1$ is written $t^+(j)$ and an opposite transition is denoted $t^-(j)$. An exact expression for the mean first passage time from zero individuals correct to $N_{ESC}$ is \cite{gardiner}

@@

\begin{align}
N \int_{0}^{x_{ESC}} &\exp\left[N\phi(x)\right] f(x) dx    \nonumber \\ 
 &\approx \sqrt{\frac{2\pi N}{\vert \phi''(x_2) \vert}}  \exp\left[N\phi(x_2)\right] f(x_2) 
\label{laplace_out}
\end{align}
As $x_1<x_2$ this may be repeated for the inner integral to leave
\begin{equation}
\tau \approx  \frac{2\pi \exp\left[N \phi(x_2)-N \phi(x_1) \right] }{t^+(x_1)\sqrt{\vert \phi''(x_2) \vert \phi''(x_1)}}.
\label{laplace_approx}
\end{equation}
Taking logarithms, derivatives, and again assuming large $N$, leaves
\begin{equation}
N^{-1}  \frac{d \ln \tau}{d \sigma} \approx -\int_{x_1}^{x_2} \frac{\partial}{\partial \sigma}\ln\left(\frac{t^+(y)}{t^-(y)}\right) dy
\label{laplace_approx}
\end{equation}

%\section{Transition rates}
The dynamics of an update step in stochastic decision process are as follows; first an individual is selected at random from the population, this individual has a personal estimate of the environmental cue $g_i$ and observes the opinions of all its neighbors. The decision rule of Eqn.~\ref{opt_rule} is then employed and the individual will adopt, or remain in, the state most likely to be correct.
For any value of social information $x_L$ there is a value of personal information above which the individual will switch to $S=1$. We denote this value as $g_c$ and from Eqn.~\ref{opt_rule} 
\begin{equation}
g_c(x_L) = \sigma\ln\chi  \left(\frac{1}{2}-x_L\right) 
\end{equation}
If $g_i>g_c$ then the individual will switch to $S=1$ otherwise $S=-1$. 
Dealing first with the transition rate toward the correct state, we may write this as
\begin{equation}
%t^+(x) = P(S_i=-1)\sum_{j=0}^{N_s}P\left(g_i>g_c(j/N_s)\right)P\left(x_L=j/N_s|x\right)
t^+(x) = P(S_i=-1)\sum_{x_{Li}}P\left(g_i>g_c(x_{Li})\right)P\left(x_{Li}|x\right)
\end{equation}
where $i$ denotes the selected individual and the summation is taken over all possible values of $x_{Li}$.
he product of the probability that the focal individual is in the state $S=-1$ by the probability of a given value of $x_L$ by the
